---
title: "FireRehabAnalyses"
author: "A. Clason"
date: "January 6, 2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(sf)
library(data.table)
library(rsq)
library(car)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(raster)
library(lme4)
library(lmerTest)
library(MuMIn)
library(soiltexture)
library("stringr")
library(xml2)
library(tidyr)
library(ggsci)
source("./R/CarbonFunctions.R")

```

Import data
```{r include=FALSE}
DatPath <- "./Inputs/"
#Plot
FR_treatments <- fread(paste0(DatPath,"FR_Treatments.csv")) # has field plot assessed treatments and fire year
#decided to change FR08 to NP because while it was a plantation, it was burned at low severity and not planted after the fire.
#Tree data:
A1trees <- fread(paste0(DatPath,"A1trees.csv"))
B1trees <- fread(paste0(DatPath,"B1trees.csv"))
Regen <- fread(paste0(DatPath,"Regen.csv"))

#Soils data:
Soils <- fread(paste0(DatPath,"Soils.csv"))

#Woody debris:
cwd <- read.csv(paste0(DatPath,"FireRehabData_CWD.csv"),header = T,stringsAsFactors = T)
fwd <- read.csv(paste0(DatPath,"FireRehabData_FWD.csv"),header = T,stringsAsFactors = T)
line <- read.csv(paste0(DatPath,"FireRehabData_TransectDistance.csv"),header = T,stringsAsFactors = T) 

#Plot treatment cleaning
FR_treatments[,ID:=as.factor(ID)][,Planted:=as.factor(Planted)][,CumBurnSev:= as.factor(CumBurnSev)]
FR_treatments$CumBurnSev <- factor(FR_treatments$CumBurnSev, levels=c("L","LM","M","MH","H"))
FR_treatments[,FIRE_YR_cat:=ifelse(FIRE_YEAR<=1979,"1960-1979",
                                   ifelse(FIRE_YEAR>1979 & FIRE_YEAR<=1989,"1980-1989",
                                          ifelse(FIRE_YEAR>1989 & FIRE_YEAR<=1999,"1990-1999",
                                                 ifelse(FIRE_YEAR>1999 & FIRE_YEAR<=2010,"2000-2009",
                                                        "2010+"))))]
FR_treatments[,TimeSinceFire := 2020 - FIRE_YEAR]

#Climate
AnnClim <- fread(paste0(DatPath,"FR_plots_Clim_Normal_1961_1990Y.csv"))
FR_treatments <- merge(FR_treatments, AnnClim[,.(ID1,MAT, MWMT, MCMT, TD, MAP, MSP, AHM, SHM, DD_0, DD5, DD_18,
                                 DD18, NFFD, bFFP, eFFP, FFP, PAS, EMT, EXT, MAR, Eref, CMD, RH,
                                CMI, DD1040)],by.x="ID",by.y="ID1")

```


## Calculating carbon pools
This code co-written by Alana Clason, Erica Lilles and Ingrid Farnell

### Tree Carbon
A1 plot = 5.64m radius = 100m2, B1 = 11.28m radius= 400 m2, A2 = 3.99m radius = 50m2
#convert to Mg/ha: Kg/m2 x 10
(Kg C in the tree)/plot are(m2) x (1Mg/1000kg) x (10000m2/ha) == (Kg C in the tree)/plot are(m2) x 10
also here, we account for the sub-sampling of some plots
```{r include=FALSE}
#FRtreatment <- fread("FR_treatments.csv")
table(A1trees[,Species])
table(A1trees[,Tree_class])
table(B1trees[,Species])
table(B1trees[,Tree_class])
A1trees[,Species := as.factor(Species)]
B1trees[,Species := as.factor(Species)]
#A1 trees biomass
t <- vector()
for(i in 1:nrow(A1trees)){
  t[i] <- TreeCarbonFN(Species = A1trees[i,Species], DBH= A1trees[i,DBH], 
                        HT= A1trees[i,Height], Tree_class=A1trees[i,Tree_class])
}
table(A1trees[,`Sub-plot`])
A1trees[,AreaSearchM2 := ifelse(`Sub-plot`=="A1",100, ifelse(`Sub-plot`=="A2",50,400))]
A1trees[Notes=="only 1/4 of A1",AreaSearchM2:=100/4]
A1trees[Notes=="only 1/2 of A1",AreaSearchM2:=100/2]
A1trees[Notes=="only 1/4 of A2",AreaSearchM2:=50/4]
A1trees[,MgPerHa_fac:=10/AreaSearchM2] #calculating the factor to mulitply to get Mg/Ha
A1trees[,CarbonPerHa:=t*MgPerHa_fac]
A1trees[is.na(CarbonPerHa)]#No A1 trees, biomass==0
A1trees[is.na(CarbonPerHa), CarbonPerHa:=0]
A1trees[CarbonPerHa ==0]

table(B1trees[,`Sub-plot`])
r <- vector()
for(i in 1:nrow(B1trees)){
  r[i] <- TreeCarbonFN(Species = B1trees[i,Species], DBH= B1trees[i,DBH], 
                        HT= B1trees[i,Height], Tree_class=B1trees[i,Tree_class])
}
B1trees[,AreaSearchM2 := ifelse(`Sub-plot`=="A1",100, ifelse(`Sub-plot`=="A2",50,400))]
B1trees[,MgPerHa_fac:=10/AreaSearchM2] #calculating the factor to mulitply to get Mg/Ha
B1trees[,CarbonPerHa:=r*MgPerHa_fac]
B1trees[is.na(CarbonPerHa)]#No B1 trees, biomass==0
B1trees[is.na(CarbonPerHa), CarbonPerHa:=0]

#merge A1 and B1 together
A1B1trees <- rbind(A1trees[,.(PlotID,`Sub-plot`,Species,DBH,Height,Tree_class,AreaSearchM2,CarbonPerHa)], 
                   B1trees[,.(PlotID,`Sub-plot`,Species,DBH,Height,Tree_class,AreaSearchM2,CarbonPerHa)])
A1B1trees[,PHF:=10000/AreaSearchM2] #accounting for smaller search areas

#calculate stubbs ((0.33*3.14)*(0.00008*((SSdat$DBH^2)*(5))*100)) or use Ung to make biomass and then cut it off:
#Live trees by plot
LiveTree_plots <- merge(FR_treatments[,.(ID)],A1B1trees[Tree_class<3, sum(CarbonPerHa),by="PlotID"],
                        by.x="ID", by.y="PlotID", all=TRUE)
LiveTree_plots$V1[is.na(LiveTree_plots$V1)] <-0
setnames(LiveTree_plots,"V1","LiveTreeCperHa")

#Dead trees by plot
DeadTree_plots <- merge(FR_treatments[,.(ID)],A1B1trees[Tree_class>=3, sum(CarbonPerHa),by="PlotID"],
                        by.x="ID", by.y="PlotID", all=TRUE)
DeadTree_plots$V1[is.na(DeadTree_plots$V1)] <-0
setnames(DeadTree_plots,"V1","DeadTreeCperHa")
live_dead_CperHa <- merge(LiveTree_plots,DeadTree_plots,by="ID")

FR_treatments <- merge(FR_treatments,live_dead_CperHa,by="ID")

```


#### Regen: 
In the "regen-sapling" if the plot said A2 these tallies were already multiplied by 2 as it says in the notes
```{r include=FALSE}
#or just do a setnames()
Regen[,Height_class:=`Height_class(cm)`][,`Height_class(cm)`:=NULL]
Regen$Diam_est <- 1 # for 31-130 cm height class
Regen$Diam_est[Regen$Height_class == "0-30"] <- 0.1
h <- vector()
for(i in 1:nrow(Regen)){
  h[i] <- RegenCarbonFN_Ung(Species = Regen[i,Species], Diam_est = 0.1, 
                            Height_class = Regen[i,Height_class], Health = Regen[i,`Live/Dead`] )
}
Regen[,AreaSearchM2:=ifelse(`Sub-Plot`=="A1",100,50)]
Regen[,MgPerHa_fac:=10/AreaSearchM2]
Regen[,CarbonPerHa := h*Tally*MgPerHa_fac] 
Regen[CarbonPerHa==0]
Regen[is.na(CarbonPerHa)]
PlotRegen <- Regen[,.(Regen_MGHa = sum(CarbonPerHa)),by=c("PlotID","Live/Dead")]
PlotRegen <- dcast(PlotRegen, PlotID ~ `Live/Dead`, value.var="Regen_MGHa")
setnames(PlotRegen,c("L","D"),c("Regen_L_MGHa","Regen_D_MGHa"))

FR_treatments <- merge(FR_treatments,PlotRegen,by.x="ID",by.y="PlotID")
```

#root biomass: - calculating this off of total live carbon below 0.22xlive stand biomas 
Wang et al 2012 decay rate of dead roots. Only applied this to dead trees, but left dead roots with "non-decayed" carbon
decided to remove dead roots from dead carbon pool
```{r}
FR_treatments[,LiveRootC:=((LiveTreeCperHa*0.222)+(Regen_L_MGHa*0.222)),by="ID"] #root =  0.222 x live biomass
FR_treatments[,DeadRootC:=((DeadTreeCperHa*exp(-0.1044*TimeSinceFire))+ #decaying the tree roots by time since fire
               (Regen_D_MGHa*0.222)),by="ID"] #root =  0.222 x  dead biomass
#Dead root function from Wang et al 2012 
#rdec <- function(year,startC){startC*exp(-0.1044*year)}
#plot(rdec(startC=3.988444e-05, year = seq(1,60,by=1)))
```

### Soil carbon
#### Mineral soil (T/ha) = volume(m3/ha)
# calculate the fine fraction BulkDensity
#fine fraction bulk density defined in Bulmer and Simpson 2010 (Soil Compaction Reduced the Growth of Lodgepole Pine and Douglas-fi r Seedlings in Raised Beds after Two Growing Seasons) Fine fraction
# "bulk density was determined after passing the sample through a 2-mm sieve and assuming a coarse fragment specifi c gravity of 2650 kg m−3 according to:
#  DBff = (Ws – Wcf )/[Vtot – (Wcf/2650)]
#where DBff is fine fraction bulk density,
#Ws is the total weight of the sample retrieved from the bulk density core,
#Wcf is the weight of coarse fragments retained on the 2-mm sieve,
#Vtot is the total volume of the sample core.
# The 500 is an estimate of the organic fragment particle density I usually just use the density of wood from a table like this (Density of wood in kg/m3, g/cm3, lb/ft3 - the ultimate guide - EngineeringClicks), from e-mail with Chuck Bulmer
#Can get a better esimate by putting roots and charcoal into graduated cylinder
#fine fraction bulk density defined in Bulmer and Simpson 2010 (Soil Compaction Reduced the Growth of Lodgepole Pine and Douglas-fi r Seedlings in Raised Beds after Two Growing Seasons) Fine fraction

# "bulk density was determined after passing the sample through a 2-mm sieve and assuming a coarse fragment specifi c gravity of 2650 kg m−3 according to:

#  DBff = (Ws – Wcf )/[Vtot – (Wcf/2650)]

#where DBff is fine fraction bulk density,

#Ws is the total weight of the sample retrieved from the bulk density core,

#Wcf is the weight of coarse fragments retained on the 2-mm sieve,

#Vtot is the total volume of the sample core.

# The 500 is an estimate of the organic fragment particle density I usually just use the density of wood from a table like this (Density of wood in kg/m3, g/cm3, lb/ft3 - the ultimate guide - EngineeringClicks), from e-mail with Chuck Bulmer

#Can get a better esimate by putting roots and charcoal into graduated cylinder
option 2 for bulk density:
BulkDens[,BDest_finefract_g_cm3 := -1.977+4.105 x (C_pro/0.47)- 1.229 x log(C_pro/0.47) - 0.103*log(C_pro/0.47)^2]

Erica changed this so roots are not added in (C in roots is estimated from tree biomass)
it would be interesting to look at the C in roots from our samples and compare with the estimates from tree biomass
```{r include=FALSE}
#Change percent C from lab analysis to proportion C
Soils[,C_pro := MinSoil_C_PC/100]

#fill in missing BD volume for now
#Soils[is.na(BulkDensity_ml)]
#Soils$BulkDensity_ml[is.na(Soils$BulkDensity_ml)==TRUE]<-mean(Soils$BulkDensity_ml, na.rm = TRUE)

#calculating bulk density
Soils[,TotalSample_wgt_kg := (MinSoilAll_DryWgt-MinSoil_TinWgt+ #this includes the coarse fragments
                                    Root_wgt_inMin+Black_C_wgt_inMin+
                                    Litter_Wgt_inMin)/1000]
Soils[,CoarseFrag_wgt_kg := (Coarse_frag_weight_W_tin - MinSoil_TinWgt)/1000]
#Bulk density calculations:
Soils[,OrgFrag_wgt_kg := (Root_wgt_inMin + Black_C_wgt_inMin + Litter_Wgt_inMin)/1000]
Soils[,FineFract_wgt_kg := (MinSoilAll_DryWgt - Coarse_frag_weight_W_tin)/1000 - OrgFrag_wgt_kg]
Soils[,CoarseAndOrgFragM3M3 := (CoarseFrag_wgt_kg/2650 + OrgFrag_wgt_kg/500)/ (BulkDensity_ml/1000000) ]

#### Fine fraction BulkDensity ####
#fine fraction bulk density defined in Bulmer and Simpson 2010 (Soil Compaction Reduced the Growth of Lodgepole Pine and Douglas-fi r Seedlings in Raised Beds after Two Growing Seasons) Fine fraction
Soils[,BDcalc_finefract := FineFract_wgt_kg/((BulkDensity_ml/1000000)-
                                               (CoarseFrag_wgt_kg/2650)-(OrgFrag_wgt_kg/500))]#kg/m3
FR_minSOC_finefract_kg_m2 <- Min_SOC(Soc = Soils[,C_pro], BD = Soils[,BDcalc_finefract],
                                     depth = Soils[,BDDepth], CoarseFrags = Soils[,CoarseAndOrgFragM3M3])

#### Soil Organic Carbon ####
#using 0.5 biomass to C conversion for litter and roots as per Michelle equations
#using a 0.75 char to C mass conversion from Donato et al. 2009 (Quantifying char in postfire woody detritus inventories)
Soils[,FR_BlackC_kg := (Black_C_wgt_inMin*0.75)/1000] #removed roots, and added litter to litter
Soils[,FR_BlackC_kg_m2:= FR_BlackC_kg/(BulkDensity_ml/1000000) * BDDepth]

Soils[,FR_SOM:= C_pro/0.47]
Soils[,BDcalc_finefract_g_cm3 := BDcalc_finefract*1000/1e+06] #calculate finefraction

#Test the predicted (Perie and Ouimet) BD vs observed BD
Soils[,BDest_finefract_g_cm3 := -1.977+4.105*FR_SOM - 1.229*log(FR_SOM) - 0.103*log(FR_SOM)^2] #estimate fine fraction
#plot(Soils[,FR_SOM],Soils[,BDcalc_finefract_g_cm3])
#x<-Soils$FR_SOM[order(Soils$FR_SOM)]
#y<- -1.977+4.105*x - 1.229*log(x) - 0.103*log(x)^2
#lines(x,y)

#plot(Soils$BDest_finefract_g_cm3,Soils$BDcalc_finefract_g_cm3)
#abline(0,1)
#### Estimating missing data in plot 63 
#use estimated fine fraction from Perie and Ouimet
Soils[ID=="FR63",BDcalc_finefract := BDest_finefract_g_cm3*1000]
#back calculate likely volume of hole for 63 to estimate other C volume
Soils[ID=="FR63", BulkDensity_ml := (FineFract_wgt_kg/(BDcalc_finefract + (CoarseFrag_wgt_kg/2650) -
                                                         (OrgFrag_wgt_kg/500)))*1000000]
Soils[ID=="FR63",FR_BlackC_kg_m2 := FR_BlackC_kg/(BulkDensity_ml/1000000) *0.2] #calculate black C Kg/m2
Soils[ID=="FR63",CoarseAndOrgFragM3M3 := (CoarseFrag_wgt_kg/2650 + OrgFrag_wgt_kg/500)/ (BulkDensity_ml/1000000) ]
FR_minSOC_finefract_kg_m2[63] <- Min_SOC(Soc = Soils[ID=="FR63",C_pro], 
                                         BD = Soils[ID=="FR63",BDcalc_finefract],
                                         depth = 0.2, 
                                         CoarseFrags = Soils[ID=="FR63",CoarseAndOrgFragM3M3])
FR_minSOC_kg_m2 <-FR_minSOC_finefract_kg_m2 + Soils[,FR_BlackC_kg_m2] #fine fraction SOC + BlackCarbon C
FR_minSOC_Mg_ha <- (FR_minSOC_kg_m2/1000)*10000 #convert to Mg per hectare
FR_treatments[,MinSoilC_Mgha := FR_minSOC_Mg_ha]

####### ADD TEXTURES ########
Soils[,CLAY := sum(MinSoil_Clay_PC_hyd,MinSoil_Clay_PC_h202_hyd,na.rm = TRUE), by="ID"]
Soils[,SILT := sum(MinSoil_Silt_PC_hyd, MinSoil_Silt_PC_h202_hyd,na.rm = TRUE), by="ID"]
Soils[,SAND := sum(MinSoil_Sand_PC_hyd,MinSoil_Sand_PC_h202_hyd,na.rm = TRUE), by="ID"]
Soils[,TotalTexture := CLAY + SILT + SAND]

#text.dat<-TT.normalise.sum(tri.data = Soils)
#text.class<-TT.points.in.classes(tri.data = Soils, text.sum = Soils$TotalTexture, 
 #                                class.sys = "CA.EN.TT", tri.sum.tst = FALSE)
#text.class<-as.data.frame(text.class)

#TT.plot( class.sys = "CA.EN.TT" )
#TT.plot(tri.data = text.dat,  class.sys = "CA.EN.TT", blr.tx = c("SAND", "CLAY", "SILT"), css.lab = c( "CLAY","SILT", "SAND" ))
#Soils[,Texture_7cl := ifelse(text.class$Cl == 1, "Clay",
 #                        ifelse(text.class$ClLo == 1, "ClayLoam",
  #                              ifelse(text.class$SiLo == 1, "SiltLoam",
   #                                    ifelse(text.class$L == 1, "Loam",
    #                                          ifelse(text.class$SaLo == 1, "SandyLoam",
     #                                                ifelse(text.class$LoSa == 1, "LoamySand", "Sand"))))))]
#Soils[,Texture_4cl := ifelse(text.class$Cl == 1, "Clay",
 #                        ifelse(text.class$ClLo == 1, "Clay",
  #                              ifelse(text.class$SiLo == 1, "SiltLoam",
   #                                    ifelse(text.class$L == 1, "Loamy",
    #                                          ifelse(text.class$SaLo == 1, "Loamy",
     #                                                ifelse(text.class$LoSa == 1, "Sand", "Sand"))))))]
#Soils[,Texture_7cl := as.factor(Texture_7cl)][,Texture_4cl := as.factor(Texture_4cl)]

#ggplot(Soils, aes(x=Texture_7cl,y=FR_minSOC_Mg_ha))+
 # geom_boxplot()

### Add C:N ratio
Soils[,CN := MinSoil_C_PC/MinSoil_N_PC]

#combine this into the plot treatments
#FR_treatments <- merge(FR_treatments, Soils[,.(ID,CN,CLAY,Texture_7cl,Texture_4cl)],by="ID")
FR_treatments <- merge(FR_treatments, Soils[,.(ID,CN,CLAY)],by="ID")
FR_treatments[,CoarseFrag_wgt_kg := Soils[,CoarseFrag_wgt_kg]]

ggplot(FR_treatments, aes(x=CLAY,y=MinSoilC_Mgha))+
  geom_point()+
  geom_smooth(method="lm")

ggplot(FR_treatments, aes(x=CoarseFrag_wgt_kg,y=MinSoilC_Mgha))+
  geom_point()

```

#### Litter
collected in 4 * 20cm x 20cm area = 0.16m2 = 1.6e-05 ha 
```{r include=FALSE, eval=TRUE, echo=FALSE}
#add litter from mineral soil to litter
FR_treatments[,Litter_C_g := Soils[,(Litter_DryWgt*(Litter_C_PC/100)) + Litter_Wgt_inMin*0.5]]
FR_treatments[,Litter_C_MgHa := Litter_C_g/16]
```

#### Forest floor
area collection = 4 x 0.20 x 0.20 = 0.16m2
```{r include=FALSE, eval=TRUE, echo=FALSE}
#calculating the volume of forest floor (F and H layers)
Soils[,ForestFloor_C_g := sum((ForFloor_DryWgt*ForFloor_C_PC/100), Wood_chunks_inFF*0.5,
       Black_C_wgt_inFF*0.75,na.rm=TRUE), by="ID"]
#tiny amt of FF in FR03 didn't make it to the lab, so just use 0.5
Soils[ID=="FR03",ForestFloor_C_g := ForFloor_DryWgt*0.5+Wood_chunks_inFF*0.5+Black_C_wgt_inFF*0.75 ]
#the only true NA (missing sample) is FR02
Soils[ID=="FR02",ForestFloor_C_g := NA]
      
FR_treatments[,ForestFloor_C_g := Soils[,ForestFloor_C_g]]
FR_treatments[,ForestFl_MgHa :=ForestFloor_C_g/16] #converting to Mg/Ha
```


### Woody debris carbon
#### Coarse woody debris already in Mg/ha = T/ha
```{r include=FALSE,eval=TRUE, echo=FALSE}
source("./R/CWD_carbon.R")
cwd.carbon.plot <- as.data.table(cwd.carbon.plot)
# Add plot description to data
# Import treatment data data
plot.desc<- FR_treatments
str(plot.desc)
# First column imported weird - rename it
names(plot.desc)[names(plot.desc) == "ID"] <- "PlotID"
# Merge dataframes via PlotID
cwd.desc <- merge(cwd.carbon.plot, plot.desc, by = "PlotID")
cwd.desc$FIRE_YR_cat <- as.factor(cwd.desc$FIRE_YR_cat)

FR_treatments[,CWD_C:=cwd.carbon.plot$carbon]

```

#### Fine woody debris (T/ha) = volume(m3/ha)
```{r include=FALSE,eval=TRUE, echo=FALSE}
source("./R/FWD_carbon.R")
fwd.carbon.plot <- as.data.table(fwd.carbon.plot)
# Add plot description to data
# Import treatment data data
plot.desc<- FR_treatments
str(plot.desc)

# First column imported weird - rename it
names(plot.desc)[names(plot.desc) == "ID"] <- "PlotID"

# Merge dataframes via PlotID
fwd.desc <- merge(fwd.carbon.plot, plot.desc, by = "PlotID") #not working now
fwd.desc <- as.data.frame(fwd.desc)
fwd.desc$FIRE_YR_cat <- as.factor(fwd.desc$FIRE_YR_cat)

FR_treatments[,FWD_C:=fwd.carbon.plot$carbon]
```



## Data setup
```{r include=FALSE,eval=TRUE, echo=FALSE}

CarbonPlots <- melt(FR_treatments, id.vars = c("ID", "Date", "Age","Planted","CumBurnSev","CumBurnSevCat",
                                               "Aspect","Slope_PC", "SlopePos","FIRE_NUMBE", "FIRE_NAME", 
                                               "FIRE_YEAR", "area","FIRE_YR_cat"),
                    measure.vars = c("LiveTreeCperHa","Regen_L_MGHa","Regen_D_MGHa","DeadTreeCperHa",
                                     "MinSoilC_Mgha","Litter_C_MgHa",
                                     "ForestFl_MgHa","CWD_C","FWD_C","LiveRootC"),
                    variable.name = "CarbonSource",
                    value.name = "CarbonMgHa")
FR_treatments <- merge(FR_treatments,CarbonPlots[,.(TotalCarbon = sum(na.omit(CarbonMgHa))), by="ID"],by="ID")
FR_treatments <- merge(FR_treatments,CarbonPlots[CarbonSource!="MinSoilC_Mgha",
                                                 .(TotalCarbon_NotMin = sum(na.omit(CarbonMgHa))), by="ID"],by="ID")
FR_treatments <- merge(FR_treatments,CarbonPlots[CarbonSource=="LiveTreeCperHa"|
                                                   CarbonSource=="DeadTreeCperHa"|
                                                   CarbonSource=="CWD_C"|
                                                   CarbonSource=="FWD_C",.(AboveGrCarbon = sum(na.omit(CarbonMgHa))),
                                                 by="ID"],by="ID")
FR_treatments <- merge(FR_treatments,CarbonPlots[CarbonSource=="MinSoilC_Mgha"|
                                                   CarbonSource=="Litter_C_MgHa"|
                                                   CarbonSource=="ForestFl_MgHa"|
                                                   CarbonSource=="LiveRootC"|
                                                   CarbonSource=="DeadRootC",
                                                 .(BelowGrCarbon = sum(na.omit(CarbonMgHa))),
                                                 by="ID"],by="ID")
FR_treatments <- merge(FR_treatments,CarbonPlots[CarbonSource=="DeadTreeCperHa"|
                                                   CarbonSource=="Regen_D_MGHa"|
                                                   CarbonSource=="CWD_C"|
                                                   CarbonSource=="FWD_C",
                                                   #CarbonSource=="DeadRootC",
                                                 .(DeadCarbon = sum(na.omit(CarbonMgHa))),
                                                 by="ID"],by="ID")
FR_treatments <- merge(FR_treatments,CarbonPlots[CarbonSource=="CWD_C"|
                                                   CarbonSource=="FWD_C",
                                                 .(WDCarbon = sum(na.omit(CarbonMgHa))),
                                                 by="ID"],by="ID")
FR_treatments <- merge(FR_treatments,CarbonPlots[CarbonSource=="LiveTreeCperHa"|
                                                   CarbonSource=="Regen_L_MGHa"|
                                                   CarbonSource=="LiveRootC",
                                                 .(LiveCarbon = sum(na.omit(CarbonMgHa))),
                                                 by="ID"],by="ID")
FR_treatments <- merge(FR_treatments,CarbonPlots[CarbonSource=="ForestFl_MgHa"|
                                                   CarbonSource=="Litter_C_MgHa",
                                                 .(ForestFloorTotC = sum(na.omit(CarbonMgHa))),
                                                 by="ID"],by="ID")
FR_treatments <- merge(FR_treatments,CarbonPlots[CarbonSource=="ForestFl_MgHa"|
                                                   CarbonSource=="MinSoilC_Mgha",
                                                 .(Min_FF = sum(na.omit(CarbonMgHa))),
                                                 by="ID"],by="ID")
FR_treatments <- merge(FR_treatments,CarbonPlots[CarbonSource=="MinSoilC_Mgha" |
                                                   CarbonSource=="ForestFl_MgHa"|
                                                   CarbonSource=="Litter_C_MgHa",
                                                 .(Min_FF_L = sum(na.omit(CarbonMgHa))),
                                                 by="ID"],by="ID")
FR_treatments[,DeadConst:=(DeadCarbon+0.01)]
FR_treatments[,FFConst:=(ForestFloorTotC+0.01)]

FR_tables <- melt(FR_treatments,id.vars = c("ID","Planted","TimeSinceFire"),
                   measure.vars = c("LiveTreeCperHa","Regen_L_MGHa","LiveRootC","DeadTreeCperHa","Regen_D_MGHa",
                                    "CWD_C","FWD_C"),
                   variable.name = "CarbonPool",
                   value.name = "CarbonMgHa")
FR_tables[,TSF_cat:=ifelse(TimeSinceFire >50,55,
                            ifelse(TimeSinceFire >40,45,
                                   ifelse(TimeSinceFire > 30, 35,
                                          ifelse(TimeSinceFire>20,25,
                                                 ifelse(TimeSinceFire>10,15,
                                                        5)))))]
FR_pool_summary <- FR_tables[,.(mn_CarbonMgHa=mean(CarbonMgHa),sd_CarbonMgHa=sd(CarbonMgHa)),
                       by=c("CarbonPool","Planted","TSF_cat")]

```



#### Data distributions
```{r eval=FALSE,echo=FALSE}
#if we need 10 x the number of data points to estimate each parameter, a model with all interactions == 20 parameters. With 75 plots, that's not enough:
#looking at the distribution of response:
hist(FR_treatments[,TotalCarbon])
hist(FR_treatments[,TotalCarbon_NotMin])
hist(FR_treatments[,LiveCarbon])
hist(FR_treatments[,WDCarbon])
hist(FR_treatments[,DeadTreeCperHa])
hist(FR_treatments[,DeadCarbon])
hist(FR_treatments[,MinSoilC_Mgha])
hist(FR_treatments[,ForestFloorTotC])
#looking at the response of residuals:
hist(resid(lm(TotalCarbon ~ TimeSinceFire + Planted + CumBurnSevCatGr, data=FR_treatments)))
#residuals actually all look good for normal error distribution. Could just fit gaussian across the board?
```


### Environmental data
#### Climate variables
```{r}
biplot(princomp(FR_treatments[,.(MAT, MWMT, MCMT, TD, MAP, MSP,SHM, NFFD, FFP, PAS, EMT, EXT, MAR, Eref, CMD, RH,
                                CMI,DD5)]))
biplot(princomp(FR_treatments[,.(FFP,CMD)]))
ggplot(FR_treatments)+
  geom_point(aes(x=CMD,y=MAT))
ggplot(FR_treatments,aes(x=CMI, y=TotalCarbon))+
  geom_point()+
  geom_smooth(formula = y~x, method = "lm")
AIC(glm(TotalCarbon~CMI, data = FR_treatments))
AIC(glm(TotalCarbon~FFP, data = FR_treatments))
AIC(glm(TotalCarbon~MAT, data = FR_treatments))
summary(glm(TotalCarbon ~ MAT+MAP+PAS+NFFD+DD5+CMD, data=FR_treatments))
cor(FR_treatments[,.(DD5,MAP)])
#centre and scale covariates
FR_treatments[,DD5 := scale(DD5, center=TRUE, scale=TRUE)]
FR_treatments[,MAP := scale(MAP, center=TRUE, scale=TRUE)]

```

#### Soil variables
```{r}
#c(CN,CoarseFrag_wgt_kg, hli, tpi, CLAY) #or coarse vol %
cor(FR_treatments[,.(CN,CoarseFrag_wgt_kg,CLAY)]) #correlation amongst productivity variables
AIC(glm(MinSoilC_Mgha ~ CLAY, data=FR_treatments))

#total carbon
AIC(glm(TotalCarbon ~ CLAY, data=FR_treatments)) #weak winner
AIC(glm(TotalCarbon ~ CN, data=FR_treatments))
AIC(glm(TotalCarbon ~ CoarseFrag_wgt_kg, data=FR_treatments))
AIC(glm(TotalCarbon ~ CN + CLAY+CoarseFrag_wgt_kg, data=FR_treatments)) #winner

#live carbon
AIC(glm(LiveCarbon ~ CLAY, data=FR_treatments))
AIC(glm(LiveCarbon ~ CN, data=FR_treatments)) #winner - AIC and explains most single variable variance
AIC(glm(LiveCarbon ~ CoarseFrag_wgt_kg, data=FR_treatments))
AIC(glm(LiveCarbon ~ MinSoilC_Mgha, data=FR_treatments))
AIC(glm(LiveCarbon ~ CN + CLAY +CoarseFrag_wgt_kg, data=FR_treatments)) #same as above

#Dead carbon
AIC(glm(DeadCarbon ~ CLAY, data=FR_treatments))
AIC(glm(DeadCarbon ~ CN, data=FR_treatments)) #weak winner
AIC(glm(DeadCarbon ~ CoarseFrag_wgt_kg, data=FR_treatments))
AIC(glm(DeadCarbon ~ CN + CLAY +CoarseFrag_wgt_kg, data=FR_treatments))

#forest floor carbon
AIC(glm(ForestFloorTotC ~ CLAY, data=FR_treatments)) #weak winner
AIC(glm(ForestFloorTotC ~ CN, data=FR_treatments)) 
AIC(glm(ForestFloorTotC ~ CoarseFrag_wgt_kg, data=FR_treatments))
AIC(glm(ForestFloorTotC ~ CN + CLAY, data=FR_treatments))

#Mineral soil carbon
AIC(glm(MinSoilC_Mgha ~ CLAY, data=FR_treatments)) #major winner
AIC(glm(MinSoilC_Mgha ~ CN, data=FR_treatments)) 
AIC(glm(MinSoilC_Mgha ~ CoarseFrag_wgt_kg, data=FR_treatments))
AIC(glm(MinSoilC_Mgha ~ CN + CLAY + CoarseFrag_wgt_kg, data=FR_treatments)) #even better winner

ggplot(FR_treatments,aes(x=MinSoilC_Mgha, y=LiveCarbon))+
  geom_point()+
  geom_smooth(formula = y~x, method = "lm")

#scale and centre continuous variables
FR_treatments[,CN:=scale(CN, center=TRUE,scale=TRUE)]
FR_treatments[,CoarseFrag_wgt_kg:=scale(CoarseFrag_wgt_kg, center=TRUE,scale=TRUE)]
FR_treatments[,CLAY:=scale(CLAY, center=TRUE,scale=TRUE)]

#check correlation with climate
cor(FR_treatments[,.(FFP,CMD,CN,CoarseFrag_wgt_kg,CLAY)]) #correlation amongst productivity variables

```

#### Site variables
```{r}
#re-download DEM and re-run topography script
hli_rast <- raster("./Inputs/Rasters/DEMhli.tif")
slope_rast <- raster("./Inputs/Rasters/DEMslope.tif")
Study_plots <- read_sf("./Inputs/FRplots.shp")
crs(hli_rast)
crs(slope_rast)
crs(Study_plots)

Plot_hli <- raster::extract(hli_rast,Study_plots,method="simple")
Plot_slope <- raster::extract(slope_rast,Study_plots,method="simple")

FR_treatments[,hli := scale(Plot_hli,center=TRUE,scale=TRUE)][,slope_ra := scale(Plot_slope,center=TRUE,scale=TRUE)] #not centred or scaled
#FR_treatments[,Slope_PC_sc := scale(Slope_PC, center=TRUE,scale=TRUE)]
```

#### Fire Severity
```{r}
#re-download CanLabS and re-run topography script
Can_dNBR <- raster("E:/Spatial Data/Fire/CanLaBS/CanLaBS_dNBR1000_v0.tif")
Study_plots <- read_sf("./Inputs/FRplots.shp")
crs(Can_dNBR)
crs(Study_plots)
#project studyplot to raster (I think that's faster)
Study_plots_tr <- st_transform(Study_plots,crs=crs(Can_dNBR))
Plot_dNBR <- raster::extract(Can_dNBR,Study_plots_tr,method="simple")
#how many are NA after 1985:
FR_treatments[,dNBR:=Plot_dNBR]
FR_treatments[FIRE_YEAR>1985,.(ID, Planted, FIRE_YEAR,dNBR)]
FR_treatments[FIRE_YEAR>1985,.N]

FR_treatments[,dNBR := ifelse(!is.na(Plot_dNBR),Plot_dNBR,ifelse(FIRE_YEAR>1985,1,Plot_dNBR))]
FR_treatments[,dNBR_nsc := dNBR]
FR_treatments[,dNBR_Class := ifelse(dNBR_nsc<=100,"Unburned",
                                    ifelse(dNBR_nsc>=100 & dNBR_nsc <270,"L",
                                           ifelse(dNBR_nsc>=270 & dNBR_nsc<440, "LM",
                                                  ifelse(dNBR_nsc>=440 & dNBR_nsc<660, "MH",
                                                         "High"))))]
FR_treatments[,dNBR := scale(dNBR, center=TRUE,scale=TRUE)]
ggplot(FR_treatments, aes(x=Planted, y=dNBR))+
  geom_boxplot()
t.test(FR_treatments[Planted=="NP"]$dNBR,FR_treatments[Planted=="P"]$dNBR)

ggplot(FR_treatments, aes(x=TimeSinceFire, y=dNBR_nsc))+
  geom_point()+
  geom_smooth(method="lm")

```


## Analysis
```{r}
########################################################### 
########## All data included #############
Mod <- list()
Mod[[1]] <- lmer(TotalCarbon ~ Planted-1 + TimeSinceFire + DD5 + MAP + CN + CoarseFrag_wgt_kg +
                   CLAY+hli + slope_ra +(1|FIRE_NAME),data=FR_treatments)
Mod[[2]] <- lmer(MinSoilC_Mgha ~ Planted-1 + TimeSinceFire  + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                  slope_ra +(1|FIRE_NAME), data=FR_treatments)
Mod[[3]] <- lmer(LiveCarbon ~ Planted-1 + TimeSinceFire + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli + 
                   slope_ra +(1|FIRE_NAME), data=FR_treatments)
Mod[[4]] <- lmer(DeadCarbon ~ Planted-1 + TimeSinceFire  + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                   slope_ra+(1|FIRE_NAME),data=FR_treatments)
Mod[[5]] <- lmer(ForestFloorTotC ~ Planted-1 + TimeSinceFire  + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                  slope_ra +(1|FIRE_NAME),data=FR_treatments)

ModNoPlant <- list()
ModNoPlant[[1]] <- lmer(TotalCarbon ~  TimeSinceFire + DD5 + MAP + CN + CoarseFrag_wgt_kg +
                   CLAY+hli + slope_ra +(1|FIRE_NAME),data=FR_treatments)
ModNoPlant[[2]] <- lmer(MinSoilC_Mgha ~ TimeSinceFire  + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                  slope_ra +(1|FIRE_NAME), data=FR_treatments)
ModNoPlant[[3]] <- lmer(LiveCarbon ~ TimeSinceFire + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli + 
                   slope_ra +(1|FIRE_NAME), data=FR_treatments)
ModNoPlant[[4]] <- lmer(DeadCarbon ~ TimeSinceFire  + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                   slope_ra+(1|FIRE_NAME),data=FR_treatments)
ModNoPlant[[5]] <- lmer(ForestFloorTotC ~ TimeSinceFire  + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                  slope_ra +(1|FIRE_NAME),data=FR_treatments)
#Total, Mineral soil, live, dead, forestfloor
lapply(Mod,AIC)
lapply(ModNoPlant,AIC)
lapply(Mod,rsq.lmm)
lapply(Mod,summary)
lapply(ModNoPlant,summary)
lapply(ModNoPlant,rsq.lmm)

########################################################### 
########## With fire severity ############
Mod <- list()
Mod[[1]] <- lmer(TotalCarbon ~ Planted-1 + TimeSinceFire  + dNBR+ DD5 + MAP + CN + CoarseFrag_wgt_kg +
                   CLAY+hli + slope_ra +(1|FIRE_NAME),data=FR_treatments)
Mod[[2]] <- lmer(MinSoilC_Mgha ~ Planted-1 + TimeSinceFire  + dNBR + DD5 + MAP + CN + CoarseFrag_wgt_kg + 
                   CLAY+hli +slope_ra +(1|FIRE_NAME), data=FR_treatments)
Mod[[3]] <- lmer(LiveCarbon ~ Planted-1 + TimeSinceFire  + dNBR+ DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli + 
                   slope_ra +(1|FIRE_NAME), data=FR_treatments)
Mod[[4]] <- lmer(DeadCarbon ~ Planted-1 + TimeSinceFire  + dNBR + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                   slope_ra+(1|FIRE_NAME),data=FR_treatments)
Mod[[5]] <- lmer(ForestFloorTotC ~ Planted-1 + TimeSinceFire + dNBR  + DD5 + MAP + CN + CoarseFrag_wgt_kg +
                   CLAY+hli + slope_ra +(1|FIRE_NAME),data=FR_treatments)

ModNoPlant <- list()
ModNoPlant[[1]] <- lmer(TotalCarbon ~  TimeSinceFire  + dNBR+ DD5 + MAP + CN + CoarseFrag_wgt_kg +
                   CLAY+hli + slope_ra +(1|FIRE_NAME),data=FR_treatments)
ModNoPlant[[2]] <- lmer(MinSoilC_Mgha ~ TimeSinceFire + dNBR + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                  slope_ra +(1|FIRE_NAME), data=FR_treatments)
ModNoPlant[[3]] <- lmer(LiveCarbon ~ TimeSinceFire + dNBR + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli + 
                   slope_ra +(1|FIRE_NAME), data=FR_treatments)
ModNoPlant[[4]] <- lmer(DeadCarbon ~ TimeSinceFire + dNBR  + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                   slope_ra+(1|FIRE_NAME),data=FR_treatments)
ModNoPlant[[5]] <- lmer(ForestFloorTotC ~ TimeSinceFire + dNBR + DD5 + MAP + CN + CoarseFrag_wgt_kg + CLAY+hli +
                  slope_ra +(1|FIRE_NAME),data=FR_treatments)

lapply(Mod,AIC)
lapply(ModNoPlant,AIC)
lapply(Mod,summary)
lapply(Mod,rsq.lmm)
lapply(ModNoPlant,summary)
lapply(ModNoPlant,rsq.lmm)
########################################################### 
#checking the effect of fire severity on forest floor
AIC(lmer(ForestFloorTotC ~ Planted-1 + TimeSinceFire + dNBR  + DD5 + MAP + CN + CoarseFrag_wgt_kg +
                   CLAY+hli + slope_ra +(1|FIRE_NAME),data=FR_treatments))
AIC(lmer(ForestFloorTotC ~ Planted-1 + TimeSinceFire + DD5 + MAP + CN + CoarseFrag_wgt_kg +
                   CLAY+hli + slope_ra +(1|FIRE_NAME),data=FR_treatments[!is.na(dNBR)]))

```

Figures
```{r}
FR_pools <- FR_treatments[,.(ID,Planted,TimeSinceFire,TotalCarbon,LiveCarbon,DeadCarbon,ForestFloorTotC,MinSoilC_Mgha)]
FR_pools[,Planted:=ifelse(Planted=="NP","Not planted", "Planted")]
FR_pools[,TSF_cat:=ifelse(TimeSinceFire >50,55,
                            ifelse(TimeSinceFire >40,45,
                                   ifelse(TimeSinceFire > 30, 35,
                                          ifelse(TimeSinceFire>20,25,
                                                 ifelse(TimeSinceFire>10,15,
                                                        5)))))]

FR_pools_m <- melt(FR_pools,id.vars = c("ID","Planted","TSF_cat"),
                   measure.vars = c("LiveCarbon","DeadCarbon","ForestFloorTotC","MinSoilC_Mgha"),
                   variable.name = "CarbonPool",
                   value.name = "CarbonMgHa")

table(FR_pools_m[,.(Planted,TSF_cat)])
FR_means <- FR_pools_m[,.(mn_CarbonMgHa=mean(CarbonMgHa),sd_CarbonMgHa=sd(CarbonMgHa)),
                       by=c("CarbonPool","Planted","TSF_cat")]

##Figure 2a
FR_pools <- FR_treatments[,.(ID,Planted,TimeSinceFire,TotalCarbon,LiveCarbon,DeadCarbon,ForestFloorTotC,MinSoilC_Mgha)]
FR_pools[,Treatment:=ifelse(Planted=="NP","Not planted", "Planted")]
FR_pools_m <- melt(FR_pools,id.vars = c("ID","Treatment","TimeSinceFire"),
                   measure.vars = c("TotalCarbon","LiveCarbon","DeadCarbon","ForestFloorTotC","MinSoilC_Mgha"),
                   variable.name = "CarbonPool",
                   value.name = "CarbonMgHa")
supp.labs <- c("Total","Live","Dead","Forest floor","Mineral soil")
names(supp.labs) <- c("TotalCarbon","LiveCarbon","DeadCarbon","ForestFloorTotC","MinSoilC_Mgha")

ggplot(FR_pools_m,aes(x=TimeSinceFire, y=CarbonMgHa))+
  geom_point(size=2,aes(colour=Treatment))+
  geom_smooth(method="lm",aes(colour=Treatment,fill=Treatment))+
  scale_color_npg()+
  scale_fill_npg()+
  theme_minimal() +
  ylab(expression("Carbon Mg" ~ ha^-1))+
  xlab("Time since fire")+
  theme(legend.position = "bottom")+
  facet_wrap("CarbonPool",labeller=labeller(CarbonPool=supp.labs))+
  theme(strip.text.x = element_text(face="bold"))
    
## Figure 2b
ggplot(FR_means)+
  geom_area(aes(x=TSF_cat,y=mn_CarbonMgHa,fill=CarbonPool))+
  facet_grid(rows="Planted")+
  scale_fill_npg(name="Carbon Pool",labels=c("Live","Dead","Forest floor","Mineral soil"))+
  scale_color_npg()+
  ylab(expression("Carbon Mg" ~ ha^-1))+
  xlab("Time since fire")+
  theme_minimal()
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),strip.background = element_blank())

######## Figure 3
Dead_part <- FR_treatments[,.(ID,Planted,TimeSinceFire, dNBR_Class,DeadTreeCperHa, Regen_D_MGHa, CWD_C, FWD_C)]
Dead_part2 <-  melt(Dead_part,id.vars = c("ID","Planted","TimeSinceFire","dNBR_Class"),
                   measure.vars = c("DeadTreeCperHa","Regen_D_MGHa","CWD_C","FWD_C"),
                   variable.name = "DeadCarbonPool",
                   value.name = "CarbonMgHa")
Dead_part2[,TSF_cat:=ifelse(TimeSinceFire >50,55,
                            ifelse(TimeSinceFire >40,45,
                                   ifelse(TimeSinceFire > 30, 35,
                                          ifelse(TimeSinceFire>20,25,
                                                 ifelse(TimeSinceFire>10,15,
                                                        5)))))]
Dead_means <- Dead_part2[,.(mn_CarbonMgHa=mean(CarbonMgHa)),by=c("DeadCarbonPool","TSF_cat","Planted")]
supp.labs <- c("Not planted","Planted")
names(supp.labs) <- c("NP","P")

### Figure 3a
ggplot(Dead_means)+
  geom_area(aes(x=TSF_cat,y=mn_CarbonMgHa,fill=DeadCarbonPool))+
  facet_grid(rows="Planted",labeller=labeller(Planted=supp.labs))+
  scale_fill_npg(name="Dead Carbon Pool",labels=c("Snags","Dead seedlings","CWD","FWD", "Dead roots"))+
  scale_color_npg()+
  ylab(expression("Carbon Mg" ~ ha^-1))+
  xlab("Time since fire")+
  theme_minimal()  

### Figure 3b
Dead_means <- Dead_part2[!is.na(dNBR_Class),.(mn_CarbonMgHa=mean(CarbonMgHa)),
                         by=c("DeadCarbonPool","dNBR_Class","Planted")]
Dead_means[,dNBR_Class:=as.factor(dNBR_Class)]
Dead_means[,dNBR_Class:=factor(dNBR_Class, levels = c("Unburned","L","LM","MH","High"))]
supp.labs <- c("Not planted","Planted")
names(supp.labs) <- c("NP","P")
ggplot(Dead_means)+
  geom_col(aes(x=dNBR_Class,y=mn_CarbonMgHa,fill=DeadCarbonPool))+
  facet_grid(rows="Planted",labeller=labeller(Planted=supp.labs))+
  scale_fill_npg(name="Dead Carbon Pool",labels=c("Snags","Dead seedlings","CWD","FWD", "Dead roots"))+
  scale_color_npg()+
  ylab(expression("Carbon Mg" ~ ha^-1))+
  xlab("Burn severity")+
  theme_minimal()  

###### Regeneration ##########
source("./R/DensityFunctions.R")
PlotRegen <- RegenDensity(Regen)
PlotRegenSum <- PlotRegen[,sum(SPH),by=c("PlotID,Species")] #combine 0-0.3 and 0.3-1.3
setnames(PlotRegenSum,"V1","SPH")
FR_Regen <- merge(FR_treatments,PlotRegenSum,by.x="ID",by.y="PlotID")
FR_Regen[,TSF_cat:=as.factor(ifelse(TimeSinceFire >50,55,
                            ifelse(TimeSinceFire >40,45,
                                   ifelse(TimeSinceFire > 30, 35,
                                          ifelse(TimeSinceFire>20,25,
                                                 ifelse(TimeSinceFire>10,15,
                                                        5))))))]
#### Figure 6
supp.labs <- c("Not planted","Planted")
names(supp.labs) <- c("NP","P")
FR_Regen[,Species:= ifelse(Species=="Lw","Larch",
                 ifelse(Species=="Ac","Cottonwood",
                        ifelse(Species=="At","Aspen",
                               ifelse(Species=="Bl","Fir",
                                      ifelse(Species =="Sx","Spruce",
                                             ifelse(Species=="Pl","Pine",
                                                    ifelse(Species=="Fd","DouglasFir",
                                                           ifelse(Species=="Ep","Birch",Species))))))))]
FR_Regen[,ConDec:=ifelse(Species=="Larch"|Species=="Fir"|Species=="Spruce"|Species=="Pine"|
                           Species=="DouglasFir","Conifer","Deciduous")]

ggplot(FR_Regen[SPH>0],aes(x=TimeSinceFire, y=SPH, color=Planted))+
  geom_point(size=4, position="jitter", alpha=0.5)+
  scale_y_log10()+
  scale_color_npg()+
  ylab(expression("Seedlings" ~ ha^-1 ~ "(log scale)"))+
  xlab("Time Since Fire")+
  facet_wrap(~Species,ncol=3)+
  #theme(panel.border=element_rect(colour="black",size=1))+
  theme_bw() +
  theme(legend.position = "bottom")+
  theme(strip.text.x = element_text(face="bold"))

#natural regeneration seedlings/ha in plots less than 30 years old:
regen_summary <- FR_Regen[TimeSinceFire<30, sum(SPH), by=c("ID","Planted","TimeSinceFire","TSF_cat")]
setnames(regen_summary,"V1","regenStems")
range(regen_summary[Planted=="P"]$regenStems)
summary(regen_summary[Planted=="P"]$regenStems)
sd(regen_summary[Planted=="P"]$regenStems)
range(regen_summary[Planted=="NP"]$regenStems)
summary(regen_summary[Planted=="NP"]$regenStems)
sd(regen_summary[Planted=="NP"]$regenStems)

#####test the difference in variance between planted and not planted:
#Young stands:
regen_summary <- FR_Regen[TimeSinceFire<30, sum(SPH), by=c("ID","Planted","TimeSinceFire","TSF_cat")]
setnames(regen_summary,"V1","regenStems")
leveneTest(regenStems ~ Planted, data = regen_summary)
t.test(regen_summary[Planted=="P"]$regenStems,regen_summary[Planted=="NP"]$regenStems)

#likely all not normal
shapiro.test(FR_regen_test[Planted=="P"]$Pl)

#using the kruskal wallis test because of large number of ties that make te Wilcox rank test difficult
wilcox.test(Pl~Planted, data=FR_regen_test, paired=FALSE)
wilcox.test(Bl~Planted, data=FR_regen_test, paired=FALSE)
wilcox.test(Ac~Planted, data=FR_regen_test, paired=FALSE)
wilcox.test(Sx~Planted, data=FR_regen_test, paired=FALSE)

ggplot(FR_Regen[SPH>0])+
  geom_boxplot(aes(x=TSF_cat,y=SPH, fill=Planted))+
  scale_color_npg()+
  scale_y_log10()+
  theme_minimal()+
  facet_wrap(~Planted)+
  theme(legend.position = "bottom")+
  theme(strip.text.x = element_text(face="bold"))

```


### Carbon sequestration over 100 years
```{r eval=FALSE}
##############################################################################
############ read in calculated Live carbon of 1ha plots #############
##############################################################################
Trees_SO <- fread("../FireRehab/Inputs/SORTIEoutputs/Trees_SO_1_75.csv") #live trees
Trees_SO <- fread("E:/SORTIE_runs/SortieExtract/SE_Trees_SO_1_75.csv") #live trees
#Trees_SO <- fread("E:/SORTIE_runs/SortieExtract/SE_Trees_SO_250.csv") #live trees

Trees_SO_live <- Trees_SO[!is.na(CarbonPerHa)]
rm(Trees_SO)
gc()
FR_Sor_trees <- Trees_SO_live[,.(LiveCperHa = sum(CarbonPerHa)),by=c("plotID","timestep")]
plotID <- c(paste0("FR0",seq(1,9,by=1)),paste0("FR",seq(10,75,by=1)))
FR_Sor_trees <- merge(FR_Sor_trees,FR_treatments[ID %in% plotID,.(ID,TimeSinceFire,Planted)],
                      by.x="plotID",by.y="ID")
FR_Sor_trees[,TSF:= timestep+TimeSinceFire]
FR_Sor_trees <- melt(FR_Sor_trees, id.vars = c("plotID", "TSF","Planted"),
                    measure.vars = c("LiveCperHa"),
                    variable.name = "CarbonPerHa",
                    value.name = "CarbonMgHa")

library(likelihood)
#Live Carbon functions 
Models <- list()
Models[[1]] <- function(parA,parB,TimeSinceFire){parA+(parB*TimeSinceFire)}   #linear function
Models[[2]] <- function(parA,parB,TimeSinceFire){(parA*TimeSinceFire)/(parB+TimeSinceFire)} #Michel-Menton function
Models[[3]] <- function(parA,parB,TimeSinceFire){parA*(1-exp(-parB^TimeSinceFire))}#Monomolecular function
Models[[4]] <- function(parA,parB,TimeSinceFire){(parA*TimeSinceFire)^2/(parB^2+TimeSinceFire^2)} #Holling type 3
Models[[5]] <- function(parA,parB,TimeSinceFire){parA+parB*log(TimeSinceFire)} #log model
Models[[6]] <- function(parA,parB,parC,TimeSinceFire){parA+(parB*TimeSinceFire)+(parC*TimeSinceFire)^2} #polynomial
Models[[7]] <- function(parA, parB, parC, TimeSinceFire){parA*(1 - exp(-1* parB * TimeSinceFire))^parC} #CR

var <- list(TimeSinceFire = "TSF",x="CarbonMgHa",mean="predicted",log=TRUE)
Planted_fits <- list()
NotPlanted_fits <- list()
for(j in 1:length(Models)){
  if(j==1|j==2|j==3|j==4|j==5){
    par <- list(parA = 0.1, parB = 0.1,sd=0.5)
    par_lo <- list(parA = -1000, parB = -1000, sd=0)
    par_hi <- list (parA = 4000, parB = 1000, sd=100)
  }else if(j==6){
    par <- list(parA = 0.1, parB = 0.1,parC=0.1, sd=0.5)
    par_lo <- list(parA = -1000, parB = -1000, parC= -1000, sd=0)
    par_hi <- list (parA = 4000, parB = 1000, parC = 1000, sd=100)
  }else{
    par <- list(parA = 0.1, parB = 0.1,parC=0.1, sd=0.5)
    par_lo <- list(parA = 0, parB = -10, parC=-10, sd=0)
    par_hi <- list (parA = 500, parB = 10, parC = 10, sd=100)
  }
  Planted_fits[[j]] <- anneal(model = Models[[j]], par = par, var = var, source_data = FR_Sor_trees[Planted=="P"],
                        par_lo=par_lo, par_hi=par_hi, pdf = dnorm, dep_var="CarbonMgHa", initial_temp = 5, 
                        temp_red = 0.85, max_iter=50000, hessian = FALSE)

  NotPlanted_fits[[j]] <- anneal(model = Models[[j]], par = par, var = var, source_data = FR_Sor_trees[Planted=="NP"],
                        par_lo=par_lo, par_hi=par_hi, pdf = dnorm, dep_var="CarbonMgHa", initial_temp = 5, 
                        temp_red = 0.85, max_iter=5000, hessian = FALSE)
}

AIC_table <- as.data.table(cbind(Model=c(paste0("Model",seq(1,length(Models)))),
                                 Planted=0,NotPlanted=0,Planted=0,NotPlanted=0,Planted=0,
                                 NotPlanted=0,Planted=0,NotPlanted=0))
for(ii in 1:length(Models)){
  if(ii <6){
    AIC_table[ii,2] <- round(Planted_fits[[ii]]$aic,0)
    AIC_table[ii,3] <- round(NotPlanted_fits[[ii]]$aic,0)
    AIC_table[ii,4] <- round(Planted_fits[[ii]]$best_pars$parA,3)
    AIC_table[ii,5] <- round(NotPlanted_fits[[ii]]$best_pars$parA,3)
    AIC_table[ii,6] <- round(Planted_fits[[ii]]$best_pars$parB,3)
    AIC_table[ii,7] <- round(NotPlanted_fits[[ii]]$best_pars$parB,3)
  }else{
    AIC_table[ii,2] <- round(Planted_fits[[ii]]$aic,0)
    AIC_table[ii,3] <- round(NotPlanted_fits[[ii]]$aic,0)
    AIC_table[ii,4] <- round(Planted_fits[[ii]]$best_pars$parA,3)
    AIC_table[ii,5] <- round(NotPlanted_fits[[ii]]$best_pars$parA,3)
    AIC_table[ii,6] <- round(Planted_fits[[ii]]$best_pars$parB,3)
    AIC_table[ii,7] <- round(NotPlanted_fits[[ii]]$best_pars$parB,3)
    AIC_table[ii,8] <- round(Planted_fits[[ii]]$best_pars$parC,3)
    AIC_table[ii,9] <- round(NotPlanted_fits[[ii]]$best_pars$parC,3)
  }
}
AIC_table
NotPlanted_fits[[1]]$R2
NotPlanted_fits[[4]]$R2
NotPlanted_fits[[5]]$R2
NotPlanted_fits[[6]]$R2
NotPlanted_fits[[7]]$R2
Planted_fits[[1]]$R2
Planted_fits[[4]]$R2
Planted_fits[[5]]$R2
Planted_fits[[6]]$R2
Planted_fits[[7]]$R2

supp.labs <- c("Not planted","Planted")
names(supp.labs) <- c("NP","P")

a <- ggplot(FR_Sor_trees[Planted=="NP"], aes(x=TSF,y=CarbonMgHa))+
      geom_point(colour="#E64B35B2", alpha = 0.1)+
      geom_line(aes(y=NotPlanted_fits[[1]]$best_pars$parA +
                      (NotPlanted_fits[[1]]$best_pars$parB*TSF)),colour="black",size=1, linetype =2)+
      geom_line(aes(y=(NotPlanted_fits[[7]]$best_pars$parA*
                         (1-exp(-1*NotPlanted_fits[[7]]$best_pars$parB*TSF))^NotPlanted_fits[[7]]$best_pars$parC)),
                colour="black",size=1, linetype=1)+
      scale_color_npg()+
      scale_fill_npg()+
      theme_minimal() +
      ylab(expression("Carbon Mg" ~ ha^-1))+
      xlab("Time since fire")+
      ylim(0,200)+
      theme(strip.text.x = element_text(face="bold"))

b <- ggplot(FR_Sor_trees[Planted=="P"], aes(x=TSF,y=CarbonMgHa))+
      geom_point(colour="#4DBBD5B2", alpha=0.1)+
      geom_line(aes(y=Planted_fits[[1]]$best_pars$parA +
                      (Planted_fits[[1]]$best_pars$parB*TSF)),colour="black",size=1, linetype=2)+
      geom_line(aes(y=(Planted_fits[[7]]$best_pars$parA*
                         (1-exp(-1*Planted_fits[[7]]$best_pars$parB*TSF))^Planted_fits[[7]]$best_pars$parC)),
                colour="black",size=1, linetype=1)+
      scale_color_npg()+
      scale_fill_npg()+
      theme_minimal() +
      ylab(expression("Carbon Mg" ~ ha^-1))+
      xlab("Time since fire")+
      ylim(0,200)+
      theme(strip.text.x = element_text(face="bold"))
ggarrange(a,b)

```


